---
title: "Lab 5"
author: "Julia Wilson"
date: "10/26/2021"
output: 
  html_document: 
    toc: yes
    theme: cerulean
    number_sections: yes
    code_folding: hide
---


```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Notes: 
- To ignore certain files, insert * = everything, and the .NAME
- code_folding: hide; where there is code that exists, the user can click if they want to see the code
- Don't have to insert #, bc number_sections does it for you
- A list: structure of data with a vector. It's a flexible way to store information. 



# Confidence Intervals 

In this section we'll learn to use the 't.test()' function as a way to return the confidence interval using the t-distribution. 

## Example 1 

Create a mock sample of observations values, and find the confidence interval. Then we'll learn how to report that. 

```{r}
pika_mass <- c(82, 94, 110, 70, 86, 99, 202)

pika_mean <- mean(pika_mass)

pika_sd <- sd(pika_mass)

pika_ci <- t.test(pika_mass)

pika_ci
```

Notes
Ways to communicate these model outputs.
You are likely to make a mistake if you write the code like this, it isn't reproducible. You want your text to update as you change data and update code. 
Hover over a function and press command return to see the outcome 
To simplify sig figs use round(data, # of sig figs)

### Risky & Tedious Way: 

Mean pika mass at Paiute Pass is 91.9 g $\pm$ 13.5 g 
(mean $\pm$ 1 standard deviation, n = 7) with a 95%
confidence interval of [79.4, 104.4] g. 

### High Initial Investment, but Safer & Better 

In-line code referencing to stored objects!

Use a single backtick on either side of a lowercase r to create
an inline bit of code. 

Mean pike mass at Paiute Pass is `r round(pika_mean, 1)` 
g $\pm$ `r round(pika_sd, 1)` g (mean $\pm$ 1 standard deviation, 
n = `r length(pika_mass)`) with a 95% confidence interval of 
[`r pika_ci$conf.int[1]`, `r pika_ci$conf.int[2]`] g.


Notes: 
- Pull things from a list in the same way you owuld refer to 
a single column in a data frame. 
- Ie my list which includes 5 vectors. If I want to pull 1 vector, 
use the $. my_list$name or my_list$age 
- If you want to see a specific element type this in console: 
pika_ci$p.value 
- CI is a vector containing two values. Now we will 'index' which
tells R what position of the vector we are interested in 
- Ie type pika_mass pulls all values in vector; then type pika_mass[#]

# Two-sample t-test

Use a two-sample t-test to test the null hypothesis that samples were drawn from populations w/ the same mean (diff in means=0).

- HO: means diff = 0 
- HA: means diff is NOT =0

This is a two-sided t-test because no directionality is implied

## Example 1

```{r}
oak_trees <- c(23,34,12,25,43,23,38)
pine_trees <- c(47,39,44,51,60,54,65)
```

Is there a significant difference in mean heights for oak and pine trees? 

```{r}
trees_t <- t.test(oak_trees, pine_trees)
```

The p_value of `r trees_t$p.value` means that .... there is `r (trees_t$p.value) *100` % chance of finding sample means at least as different as those I found by random chance if they were drawn from populations with the same mean height. 


Notes: 
- type in t.test(NAME) in console to find p value
- reject p value if 
- accept p value 
- We are creating separate vectors 
- unique(mpg$class) - shows all teh names in a column

## Example 2

Using the `mpg` data set to compare city gas mileage for SUVs and compact cars

```{r}
# Create a subsest called compact_suv that only contains observations where class is suv or compact

compact_suv <- mpg %>% 
  filter(class %in% c("suv", "compact"))
```

```{r}
ggplot(data = compact_suv, aes(x = cty)) +
  geom_histogram(bins = 12) +
  facet_wrap(~ class)

ggplot(data = compact_suv, aes(sample = cty)) +
  geom_qq() +
  facet_wrap(~ class)
```

Notes: 
Don't assume exponential. Don't let tail ends determine overall relationship. 
QQ plots should only be used in correlation with histogram 


```{r}
car_stats <- compact_suv %>% 
  group_by(class) %>% 
  summarize(mean_city = mean(cty),
            sd_cty = sd(cty),
            sample_size = n())

car_stats
```

```{r}
compact_sample <- compact_suv %>% 
  filter(class == "compact") %>%
  pull(cty)
  
suv_sample <- compact_suv %>% 
  filter(class == "suv") %>% 
  pull(cty)

cars_t <- t.test(compact_sample, suv_sample)
```

-These are two vectors you can run your p value on. 
-run cars_t in console to find p-value 
-Reject the null hypothesis of equal mean city gas mileages
-How to report findings: 
-Mean gas mileage for compact cars and suvs differs significantly (t(`r cars_t$parameter`) = `r cars_t$statistic`, p < 0.001, $\alpha$ = -.05). 








